Thermoeconomic Analysis of Biodiesel Production from Used Cooking Oils

### 2015-05-23 paper.tex
Final Publishing Version

%=================================================================
\setcounter{page}{1}
\lastpage{x}
\doinum{10.3390/------}
\pubvolume{7}
\pubyear{2015}
%\externaleditor{Academic Editor: xx}
\history{Received: 5 March 2015 / Accepted: 12 May 2015 / Published: xx May 2015}
% Add packages and commands to include here
% The amsmath, amsthm, amssymb, hyperref, caption, float and color packages are loaded by the MDPI class.
\usepackage{graphicx,soul}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{tabulary}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{rotating}
\usepackage{textcomp}
\usepackage{cleveref}
\usepackage[version=3]{mhchem}
\usepackage{subfigure}
\usepackage{bigstrut}
\usepackage{placeins}
\usepackage{mathtools}

@@

% Authors (Add full first names)
\Author{Emilio Font de Mora * and C\'esar Torres and Antonio Valero}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
 Centre of Research for Energy Resources and Consumption -- CIRCE, University of Zaragoza, Mariano Esquillor, \hl{15 50018,} Zaragoza, Spain; E-Mails: ctorrescuadra@gmail.com (C.T.);\\ valero@unizar.es (A.V.)
}

\corres{E-Mail: emilio.fontdemora@gmail.com; \linebreak \hl{Tel.: +1-111-111-111; Fax:+1-111-111-111.}\\

@@

The weakness in both methodologies is related to the assumptions made in terms of UCO losses in collection and recycling. The bottom-up approach also adds uncertainty to the data obtained per person/year and the extrapolation of these data to an entire region. As an example, the quantities obtained in the Intelligent Energy--Europe (IEE) project \hl{OILECO} %please define 
 range from 4 to 18 kg/person/year \cite{OILECO2013} in EU countries (Italy, Spain, Hungary, Slovakia and Bulgaria).

Being this the situation, some projects have estimated the potential biodiesel production in the EU. The IEE project \hl{RecOil} %please define
 states that biodiesel produced from UCO could replace 1.5\% of the EU-27 diesel consumption \cite{RecOil2013}. The final report of the IEE project \hl{BioDieNet} %please define
 \cite{BioDieNet2009} states that the total amount of recoverable UCO available in the EU-27 in 2009 was about 3.5 million tons (3950 million liters). As the EU-27 diesel consumption was 214 billion liters in 2007, around 1.8\% of the EU-27 diesel consumption could have potentially been replaced with biodiesel produced from UCO.

@@

The collection can be organized in different ways. In the HORECA sector and food industry, normally, the collection is done on-site by the authorized agent in medium-to-big containers that are transported via van or truck. In the case of households, there are different possibilities: door-to-door collection or distributed collection at points throughout an area with well-identified containers. \hl{These containers can be adapted to receive UCO in receptacles} %please check the correction
 (such as used bricks or bottles or reusable bottles designed for the purpose) or have a hole through which the oil can be poured. According to the results of IEE projects OILECO and RecOil, the best practices, showing higher acceptance, are those having collection points with \hl{containers adapted to leaving receptacles} %please check the correction
 \cite{OILECO2013,RecOil2013}. This system can also be the most sustainable if the transport of the oil to the collection point by the household consumer is not accounted for in the life \mbox{cycle assessment.}

@@

As far as the transesterification is concerned, the UCO normally follows a different process compared to the case of crude vegetable oils, which affects the consumption of materials and energy sources. The main differences are:
\begin{itemize}
\vspace{6pt}
\item The pre-treatment is not so elaborated, because the oil comes already ``clean'' from the recycling unit. Normally, cleaning is only done via centrifugation and drying to eliminate possible small particles and excess water.
\item Afterwards, the oil is esterified with an acid catalyzer to transform into esters the high quantity of free fatty acids that are normally present in UCO; this esterification is followed by a basic catalyzed transesterification. Some plants have the esterification and transesterification tanks in parallel. They treat in the esterification plant those oils having high FFA (more than 4\%) and in the transesterification plant those oils with less FFA \cite{CIEMAT2005}.
\item The ester obtained follows a cleaning and drying process. Other plants have opted to distillate it to ensure the fulfillment of the quality standards, EN 14214 \cite{CENEN14214} in the EU.
\item The glycerol does not follow an after-treatment process. The crude glycerol is either treated as a residue or sold as a low added-value co-product to soap \hl{unfolders} %please check the term; cannot verify
 or to be used as fuel in cement kilns \cite{FontdeMoraThesis2013}. Some plants have after-treatment processes similar to the ones using vegetable oils to obtain a high-quality glycerin product; these after-treatment processes normally consist of adding an acid, e.g., HCl, to split the soaps into FFA and salts, which can then be separated by decantation, and distilling the remaining flow up to \degreeC{130} to evaporate the remaining methanol \cite{FontdeMoraThesis2013}.
\vspace{6pt}
\end{itemize}

@@

Combining Equations (\ref{eq1}) and (\ref{eq2}), the exergy cost can be calculated by solving the following set of linear equations:
\begin{equation}
C_{P,i}=C_{0i}+\sum_{j=1}^{n}{y_{ji}C_{P,j}} \qquad i=1, ..., n
\end{equation}

@@

The values obtained for the UCO biodiesel can be compared to the ones obtained for biodiesels from vegetable oils in Figure \ref{fig3}. The values for vegetable oils were calculated in \cite{FontdeMoraThesis2013}. As can be seen, biodiesel production from UCO has the highest ExROI value and the second highest renewability factor, only exceeded by palm oil-based biodiesel. The high renewability of palm oil is due to the use of biomass resources to cover part of the energy needs in the cycle. In addition, UCO biodiesel has the lowest non-renewable exergy cost, which means that, overall, it consumes less non-renewable resources, \textit{i.e.}, mainly fossil fuels, than its counterparts. The left graphic, in Figure \ref{fig3}, shows also the renewable and non-renewable exergy cost per produced unit of the different biodiesels analyzed in the paper. This analysis could be extended to cover also non-edible vegetable oils, for example castor and Jatropha oils, and their blends with \hl{UCO} \cite{Kannan2013}. %where is the reference? please confirm.

@@

This paper applies the methodology and concepts introduced in \cite{FontdeMora2012,FontdeMora2013,Torres2012b} to the life cycle production of biodiesel produced from used cooking oils. Used cooking oils have the potential to replace 1.5\% to 1.8\% of the EU-27 diesel consumption with a GHG emission reduction potential of 88\% compared to fossil diesel. Given the importance that this alternative fuel is gaining and the support that is given by the energy and transport policy frameworks worldwide, it is important to assess its life cycle from the thermoeconomic perspective, analyze how the process can be improved and compare the results with other biodiesel types. The results show that biodiesel from used cooking oil is the most sustainable biodiesel from the point of view of the use of non-renewable resources. The ExROI value is higher than three, which is, according to Hall \textit{et al.} \cite{Hall2009}, the minimum \hl{EROI} %do you mean ExROI? please check or define
 value that society must attain from its energy exploitation to support continued economic activity and social function. The most important measure to improve the production cycle is substituting the fossil-derived methanol by an alcohol of biological origin, which provides an ExROI increase of 75\%.

\acknowledgments{Acknowledgments}

\hl{Please add ``Acknowledgments'' here.}

\authorcontributions{Author Contributions}
\hl{
Please state the individual contribution of each co-author to the reported research and writing of the paper. e.g., who designed research; who performed research and analyzed the data; who wrote the paper. All authors have read and approved the final manuscript.}

@@

\hl{The Authors declare no conflicts of interest.}
%please confirm ``The Authors declare no conflicts of interest''

%=================================================================
\break
\makeatletter
\renewcommand\@biblabel[1]{#1.}
\makeatother
\bibliography{Turriano}
\bibliographystyle{mdpi}

\bibliographystyle{mdpi}

@@

\bibitem[{LIFE Programme}(2012)]{LifeECOBUS}
{LIFE Programme}.
\newblock \textit{Ecobus - Collecting Used Cooking Oils to Their Recycling as Biofuel
 for Diesel Engines};
\newblock Technical Report; European Commission: Brussels, Belgium, 2012.

@@

\bibitem[EBTP2011()]{EBTP2011}
Policy toolkit to improve the support mechanisms for advanced biofuels. Availabe online: \hl{Website link } (accessed on \hl{date month year}).

\bibitem[FEDIOL(2013)]{Fediol2013}
FEDIOL. 
\newblock Vegetable oils production, import, exports and consumption 2011
 annual statistics. Available online: \hl{Website link } (accessed on \hl{date month year}).

\bibitem[{United States Department of Agriculture Foreign Agricultural
 Service}()]{USDAonline}
{United States Department of Agriculture Foreign Agricultural Service (USDA)}.
\newblock PDS - Production. Available online: \hl{Website link } (accessed on \hl{date month year}).

@@

\bibitem[Istv\'an(2011)]{OILECO2013}
Istv\'an, Z.
\newblock Analyses of the State of the Art, Main Barriers and Market Potential
 of UCOs. Available online: \hl{Website link } (accessed on \hl{date month year}).

@@

\bibitem[Lechon \em{et~al.}(2005)Lechon, Cabal, De~la Rua, Lago, Izquierdo,
 Saez, and Fdez. San~Miguel]{CIEMAT2005}
Lechon, Y.; Cabal, H.; de~la Rua, C.; Lago, C.; Izquierdo, L.; Saez, R.M.;
 Fdez. San~Miguel, M.
\newblock \textit{An\'alisis de ciclo de vida de Combustibles Alternativos para el
Transporte. Fase I. An\'alisis de ciclo de vida Comparativo del Biodiesel de
Cereales y de la Gasolina
};
\newblock Technical Report; Ministerio de Educaci\'on y Ciencia y Ministerio de
 Medio Ambiente: \hl{City, Country,} 2005.

\bibitem[IDAE(2011)]{IDAEPER}
Instituto para la Diversificación y Ahorro de la Energ\'ia (IDAE).
\newblock {\it Plan de Energ\'ias Renovables (PER) 2011--2020}. IDAE: Madrid, Spain, 2011.

@@

>>>>

%=================================================================
\setcounter{page}{6321}
\lastpage{6335}
\doinum{10.3390/su7056321}
\pubvolume{7}
\pubyear{2015}
%\externaleditor{Academic Editor: xx}
\history{Received: 5 March 2015 / Accepted: 12 May 2015 / Published: xx May 2015}
% Add packages and commands to include here
% The amsmath, amsthm, amssymb, hyperref, caption, float and color packages are loaded by the MDPI class.
\usepackage{graphicx,soul}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{tabulary}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{rotating}
\usepackage{textcomp}
\usepackage{cleveref}
\usepackage[version=3]{mhchem}
\usepackage{subfigure}
\usepackage{bigstrut}
\usepackage{placeins}
\usepackage{mathtools}

@@

% Authors (Add full first names)
\Author{Emilio Font de Mora *, C\'esar Torres and Antonio Valero}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
 Centre of Research for Energy Resources and Consumption -- CIRCE, University of Zaragoza, Mariano Esquillor 15, 50018 Zaragoza, Spain; E-Mails: ctorrescuadra@gmail.com (C.T.);\\ valero@unizar.es (A.V.)
}

\corres{E-Mail: emilio.fontdemora@gmail.com; \linebreak Tel.: +34-976-76-18-63; Fax:+34-976-73-20-78.\\

@@

The weakness in both methodologies is related to the assumptions made in terms of UCO losses in collection and recycling. The bottom-up approach also adds uncertainty to the data obtained per person/year and the extrapolation of these data to an entire region. As an example, the quantities obtained in the Intelligent Energy--Europe (IEE) project OILECO range from 4 to 18 kg/person/year \cite{OILECO2013} in EU countries (Italy, Spain, Hungary, Slovakia and Bulgaria).

Being this the situation, some projects have estimated the potential biodiesel production in the EU. The IEE project RecOil states that biodiesel produced from UCO could replace 1.5\% of the EU-27 diesel consumption \cite{RecOil2013}. The final report of the IEE project BioDieNet \cite{BioDieNet2009} states that the total amount of recoverable UCO available in the EU-27 in 2009 was about 3.5 million tons (3950 million liters). As the EU-27 diesel consumption was 214 billion liters in 2007, around 1.8\% of the EU-27 diesel consumption could have potentially been replaced with biodiesel produced from UCO.

@@

The collection can be organized in different ways. In the HORECA sector and food industry, normally, the collection is done on-site by the authorized agent in medium-to-big containers that are transported via van or truck. In the case of households, there are different possibilities: door-to-door collection or distributed collection at points throughout an area with well-identified containers. These containers can be adapted to receive UCO in used bricks, bottles or reusable bottles designed for this purpose, or have a hole through which the oil can be poured. According to the results of IEE projects OILECO and RecOil, the best practices, showing higher acceptance, are those having collection points with adapted containers~\cite{OILECO2013,RecOil2013}. This system can also be the most sustainable if the transport of the oil to the collection point by the household consumer is not accounted for in the life \mbox{cycle assessment.}

@@

As far as the transesterification is concerned, the UCO normally follows a different process compared to the case of crude vegetable oils, which affects the consumption of materials and energy sources. The main differences are:
\begin{itemize}
\vspace{6pt}
\item The pre-treatment is not so elaborated, because the oil comes already ``clean'' from the recycling unit. Normally, cleaning is only done via centrifugation and drying to eliminate possible small particles and excess water.
\item Afterwards, the oil is esterified with an acid catalyzer to transform into esters the high quantity of free fatty acids that are normally present in UCO; this esterification is followed by a basic catalyzed transesterification. Some plants have the esterification and transesterification tanks in parallel. They treat in the esterification plant those oils having high FFA (more than 4\%) and in the transesterification plant those oils with less FFA \cite{CIEMAT2005}.
\item The ester obtained follows a cleaning and drying process. Other plants have opted to distillate it to ensure the fulfillment of the quality standards, EN 14214 \cite{CENEN14214} in the EU.
\item The glycerol does not follow an after-treatment process. The crude glycerol is either treated as a residue or sold as a low added-value co-product or to be used as fuel in cement kilns \cite{FontdeMoraThesis2013}. Some plants have after-treatment processes similar to the ones using vegetable oils to obtain a high-quality glycerin product; these after-treatment processes normally consist of adding an acid, e.g., HCl, to split the soaps into FFA and salts, which can then be separated by decantation, and distilling the remaining flow up to \degreeC{130} to evaporate the remaining methanol \cite{FontdeMoraThesis2013}.
\vspace{6pt}
\end{itemize}

@@

Combining Equations (\ref{eq1}) and (\ref{eq2}), the exergy cost can be calculated by solving the following set of linear equations:
\begin{equation}
C_{P,i}=C_{0i}+\sum_{j=1}^{n}{y_{ji}C_{P,j}} \qquad i=1,\ldots, n
\end{equation}

@@

The values obtained for the UCO biodiesel can be compared to the ones obtained for biodiesels from vegetable oils in Figure \ref{fig3}. The values for vegetable oils were calculated in \cite{FontdeMoraThesis2013}. As can be seen, biodiesel production from UCO has the highest ExROI value and the second highest renewability factor, only exceeded by palm oil-based biodiesel. The high renewability of palm oil is due to the use of biomass resources to cover part of the energy needs in the cycle. In addition, UCO biodiesel has the lowest non-renewable exergy cost, which means that, overall, it consumes less non-renewable resources, \textit{i.e.}, mainly fossil fuels, than its counterparts. The left graphic, in Figure \ref{fig3}, shows also the renewable and non-renewable exergy cost per produced unit of the different biodiesels analyzed in the paper. This analysis could be extended to cover also non-edible vegetable oils, for example castor and Jatropha oils, and their blends with UCO \cite{Kannan2013}.

@@

This paper applies the methodology and concepts introduced in \cite{FontdeMora2012,FontdeMora2013,Torres2012b} to the life cycle production of biodiesel produced from used cooking oils. Used cooking oils have the potential to replace 1.5\% to 1.8\% of the EU-27 diesel consumption with a GHG emission reduction potential of 88\% compared to fossil diesel. Given the importance that this alternative fuel is gaining and the support that is given by the energy and transport policy frameworks worldwide, it is important to assess its life cycle from the thermoeconomic perspective, analyze how the process can be improved and compare the results with other biodiesel types. The results show that biodiesel from used cooking oil is the most sustainable biodiesel from the point of view of the use of non-renewable resources. The ExROI value is higher than three, which is, according to Hall \textit{et al.} \cite{Hall2009}, the minimum ExROI value that society must attain from its energy exploitation to support continued economic activity and social function. The most important measure to improve the production cycle is substituting the fossil-derived methanol by an alcohol of biological origin, which provides an ExROI increase of 75\%.

\authorcontributions{Author Contributions}
Emilio Font de Mora designed the research and written the sections related to UCO-Based Biodiesel. Thermoeconomic Analysis is written by C\'esar Torres. Antonio Valero reviewed the research. All authors have read and approved the final manuscript.

@@

The authors declare no conflict of interest.

%=================================================================
\break
\makeatletter
\renewcommand\@biblabel[1]{#1.}
\makeatother

@@

\bibitem[{LIFE Programme}(2012)]{LifeECOBUS}
{LIFE Programme}.
\newblock \textit{Ecobus---Collecting Used Cooking Oils to Their Recycling as Biofuel
 for Diesel Engines};
\newblock Technical Report; European Commission: Brussels, Belgium, 2012.

@@

\bibitem[EBTP2011()]{EBTP2011}
Policy toolkit to improve the support mechanisms for advanced biofuels. Availabe online: http://biofuelstp.eu/policy/policy-toolkit-ebtp-06-10-2011.pdf (accessed on 14 May 2015).

\bibitem[FEDIOL(2013)]{Fediol2013}
FEDIOL. 
\newblock Vegetable oils production, import, exports and consumption 2011
 annual statistics. Available online: http://www.fediol.be/web/statistics2011/1011306087/list1187970179/f1.html (accessed on 14 May 2015).

\bibitem[{United States Department of Agriculture Foreign Agricultural
 Service}()]{USDAonline}
{United States Department of Agriculture Foreign Agricultural Service (USDA)}.
\newblock PDS - Production. Available online: https://apps.fas.usda.gov/psdonline/ (accessed on 14 May 2015).

@@

\bibitem[Istv\'an(2011)]{OILECO2013}
{OilECO Project}
\newblock Project OILECO, supported by the Intelligent Energy -- Europe (IEE) Programme of the European Commision, 2013. Available online: http:/www.oileco.org (accessed on 15 \linebreak May 2015).

@@

\bibitem[Lechon \em{et~al.}(2005)Lechon, Cabal, De~la Rua, Lago, Izquierdo,
 Saez, and Fdez. San~Miguel]{CIEMAT2005}
Lechon, Y.; Cabal, H.; de~la Rua, C.; Lago, C.; Izquierdo, L.; Saez, R.M.; Fdez. San~Miguel, M.
\newblock \textit{An\'alisis de ciclo de vida de Combustibles Alternativos para el Transporte. Fase I. An\'alisis de ciclo de vida Comparativo del Biodiesel de Cereales y de la Gasolina
};
\newblock Technical Report; Ministerio de Educaci\'on y Ciencia y Ministerio de Medio Ambiente: Madrid, Spain, 2005.

\bibitem[IDAE(2011)]{IDAEPER}
Instituto para la Diversificaci\'on y Ahorro de la Energ\'ia (IDAE).
\newblock {\it Plan de Energ\'ias Renovables (PER) 2011--2020}; IDAE: Madrid, Spain, 2011.

@@

\bibitem[Sankaranarayanan and Kannan(2013)]{Kannan2013}
Sankaranarayanan, S.; Kannan, S.
\newblock Sonochemistry: A simple and efficient route for biodiesel production
  from neat and blended vegetable oils using LDH-derived base catalysts.
\newblock {\em Int. J. Adv. Eng. Sci. Appl. Math.} {\bf 2013}, {\em 5},~210--218.
